Stock forecasting algorithms and their usage

### 2015-06-18 ALR_Paper.tex
added ALR-Paper with first chapter

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[conference]{IEEEtran}
\usepackage[]{blindtext, graphicx}
\usepackage[hidelinks]{hyperref}
\graphicspath{ {Figures/} }
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{color}
\usepackage{fancyhdr}
\pagestyle{plain}

@@

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Marc Misoch}
\IEEEauthorblockA{Institute of Computer Science\\University of Applied Science\\
Mannheim, Germany\\
Email: marc.misoch@stud.hs-mannheim.de}
\and
\IEEEauthorblockN{David Marquant}
\IEEEauthorblockA{Institute of Computer Science\\University of Applied Science\\
Mannheim, Germany\\
Email: david.marquant@stud.hs-mannheim.de}
}

@@

\begin{abstract}
%\boldmath
In this paper, we propose several methods to predict future stock prices. There are
a lot of algorithms out there but some are more reliable then others. As expected,
in most cases the general rule: \textit{"the more complex, the better the prediction" } is true.
However there are some relatively easy algorithms that can predict the stock prices of the future quite well.
We are going to present and analyse some of these algorithms and take a deeper look into them.
Just to make this clear, there is not a single algorithm, that can predict the future of a 
stock 100\% correctly. The relationships of the global stock market are to complex to bring them down in a 
single computer program. We will show the experimental results we got with several algorithms and compare them.
Also we will take a short look into investment strategies based on these forecasts and take a look at the results.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
stock, forecast, prediction, paper, algorithms.
\end{IEEEkeywords}

@@

\section{Introduction}
\label{introduction}
\subsection{Introduction to stock forecasting}

There is a huge variety of methods that are used in forecasting stock prices. Many of the algorithms that have been made up just use structured data like tables with historic stock quotes in it. The aim of all these algorithms is to maximize the profit while keeping the risk as low as possible.

Since the stock market is global, there are many relationships between stocks. Some of them are paired with others so if the price of one stock goes down, the other stock will follow this trend.
Stocks are also influenced by numerous factors and these factors also have relationships between each other. This leads us to the thought, that a good stock prediction has to consider more than just past stock prices. However it is still possible to make reasonably good predictions based on just previous prices.

One very important thing is to observe the international newscasts. If the global economy, or even the economy of a single country is experiencing a financial crisis, the stock market will follow that trend. But those trends are most likely to be relatively slow and easy to predict even looking at the stock prices itself. The thing that is more dangerous are catastrophes. After a catastrophe there is often a low point on the stock market. This low point comes so fast, that only algorithms that observe the news have a chance to prevent huge losses.

Another thing to note is, that we did all our tests on historic data of the stock market. Even if we \textit{bought} a stock at some specific point of time, we did not really buy this stock on the real market. So our tests did not affect the stock market at all. Because buying and selling stocks on the market while testing, would always affect the received results.

@@

In this paper we are going to compare some algorithms and their functionality. The main goal of these algorithms is, to predict a rise or fall in the price before the actual rise or fall happens. It is very important to find some kind of pattern in the stock prices of the past to do a accurate prediction of the future prices. We also like to share some of our experimental results we had with the algorithms. We only used algorithms that do not take global news in account because we wanted to see, how good the algorithms were in an isolated environment analysing past prices only. We decided to do it that way, because the algorithms themselves should work properly and examining news would be an easy addition to these working algorithms.

The main goal of the project was, to find a good working stock forecast algorithm to predict future stock trends. Later in this paper, we will introduce parameters and variables, that give the best results combined with our algorithms.

\subsection{Document overview}
Section \ref{essentials} briefly covers the background information needed for the reader to understand the most important points about our project.

Section \ref{algorithm} describes the algorithms used in more detail.

Chapter \ref{results} is about the test-results received using the algorithms described in chapter \ref{algorithm}. All algorithms were tested with different inputs and parameters.

In section \ref{conclusion} the results are interpreted and we propose how our algorithms can be improved to be successful in the real stock market.

In chapter \ref{futurework} some thing extensions to our algorithms are given, which could be done in future.

>>>>

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[conference]{IEEEtran}
\usepackage[]{blindtext, graphicx}
\usepackage[hidelinks]{hyperref}
\graphicspath{ {Figures/} }
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{color}
\usepackage{fancyhdr}
\usepackage[utf8]{inputenc}
\pagestyle{plain}

@@

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Marc Misoch}
\IEEEauthorblockA{Fakultät für Informatik\\Hochschule Mannheim, Deutschland\\
Email: marc.misoch@stud.hs-mannheim.de}
\and
\IEEEauthorblockN{David Marquant}
\IEEEauthorblockA{Fakultät für Informatik\\Hochschule Mannheim, Deutschland\\
Email: david.marquant@stud.hs-mannheim.de}
}

@@

\begin{abstract}
%\boldmath
In diesem Paper geht es darum, wie man Börsenkurse vorhersagen kann. Dies ist kein leichtes Unterfangen, denn die Wertpapierbörse ist ein hoch komplexes Gebilde mit unglaublich vielen Abhängigkeiten. All diese Abhängigkeiten ausfindig zu machen ist unmöglich. Deshalb muss man sich mit anderen Methoden behelfen. Die von uns getestete und implementierten Algorithmen benutzen ausschließlich alten Börsenkurse um vorhersagen zu treffen. Wir haben uns für diese Methode entschieden, da wir einen gut funktionierenden Algorithmus finden wollten ohne andere Variablen wie Nachrichten mit einzubeziehen, da man funktionierende Algorithmen in dieser Hinsicht relativ leicht erweitern könnte.\\
Im Verlauf des Papers werden wir genauer auf die verschiedenen Algorithmen eingehen und diese im Detail erläutern. Außerdem werden wir unsere Testergebnisse vorstellen und versuchen möglichst gute Schlüsse aus diesen zu ziehen. Um eine Sache schon einmal vorweg zu nehmen, es gibt weltweit keinen Algorithmus der Börsenkurse zu 100\% vorhersagen kann. Dennoch ist es möglich durch hochkomplexe und geheime Algorithmen eine Genauigkeit von 60-65\% zu erreichen.  
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Wertpapierbörse, Vorhersage, Algorithmus.
\end{IEEEkeywords}

@@

\section{Einleitung}
\label{introduction}
\subsection{Einführung in die Aktienkursvorhersage}

Es gibt eine große Vielfalt an Methoden, die dafür genutzt werden Börsenkurse möglichst genau vorherzusagen. Viele dieser Algorithmen beruhen ausschließlich auf historischen Preisen der Aktie die vorhergesagt werden soll. Selbst wenn man nur diese Daten berücksichtigt kann man gute Ergebnisse erzielen. Jedoch sind diese Ergebnisse sehr anfällig gegen Kurzfristige Veränderungen durch unvorhersehbare Umwelteinflüsse.\\
Dennoch verfolgen all diese Algorithmen ein Hauptziel: \textbf{Den Profit zu maximieren während das Risiko so niedrig wie möglich gehalten wird}.

Da der Aktienmarkt global ist, gibt es viele Beziehungen zwischen den einzelnen Börsen weltweit und somit auch zwischen den einzelnen Aktien. Manche Aktienkurse hängen zusammen und werden stark von einander beeinflusst. Dies kann man an folgendem Beispiel leicht erläutern:\\
Der Aktienkurse eines Motorenherstellers verliert an Wert wenn der Aktienkurs der Autoherstellers, den er beliefert, an Wert verliert. Die meisten dieser Abhängigkeiten sind Wechselseitig und sehr stark. Dennoch sind nicht alle dieser Abhängigkeiten auf den ersten Blick so klar erkennbar. Selbst zum heutigen Zeitpunkt ist nicht immer klar wie sich der globale Markt verhalten wird, wenn es zu großen Veränderungen an einer einzigen Aktie kommt. Das lässt den Schluss zu, dass es für eine möglichst genaue und robuste vorhersage mehr braucht als nur historische Aktienkurse. Dennoch ist es möglichst selbst mit diesen Daten gute Vorhersagen zu machen.

Eine weitere wichtige Sache die man im Auge behalten muss sind die Nachrichten. Wenn die globale Wirtschaft einen negativen Trend zeigt, ist davon auszugehen, dass der Aktienmarkt diesem Trend folgt und auch einen Abstieg zu verzeichnen hat. Jedoch sind diese Trends eher langsam und man kann sie auch mit Algorithmen, deren vorhersagen auf alten Aktienkursen basieren, relativ gut vorhersehen. Eine Sache die man nicht anhand von alten Börsenkursen vorhersagen kann, sind Katastrophen. Dabei macht es wenig bis keinen Unterschied ob es sich um Naturkatastrophen oder vom Menschen verursache Katastrophen handelt. Falls so ein Fall eintritt kann der globale Aktienmarkt innerhalb von wenigen Minuten einen enormen Absturz erfahren. Das führt dazu, dass eine wirklich sichere Vorhersage nur mit Algorithmen möglich ist, die auch die Nachrichten nach Neuigkeiten durchsuchen, die den Aktienmark betreffen werden.

All unsere Test wurden mit echten historischen Aktienkursen durchgeführt. Dennoch haben wir nicht an der Wertpapierbörse gehandelt. Die Aktien mit denen wir gearbeitet haben wurden von uns nie reell an der Börse gekauft und hatten somit hatten unsere Tests auch keinen Einfluss auf den Aktienmarkt.

@@

Dieses Paper dient dazu, verschieden Algorithmen und ihre Arbeitsweise zu vergleichen. Dabei werden wir sowohl auf die Implementierung selbst als auch auf die Überlegungen die dahinter stehen eingehen. Das Hauptziel all dieser Algorithmen ist das selbe. Es soll ein Anstieg oder Abfall eines Aktienkurses vorhergesagt werden, noch bevor dieser passiert. Dafür ist es sehr wichtig ein geeignetes Muster zu finden, mit dessen Hilfe man von alten Aktienkurse auf zukünftige schließen kann. Im Anschluss werden wir auf die Test eingehen die wir mit den Implementierten Algorithmen durchgeführt haben. Die Algorithmen die wir in diesem Paper ausführen möchten, benutzen nur Aktienpreise der Vergangenheit und versuchen mit deren Hilfe eine möglichst genaue Vorhersage zu treffen. Dafür haben wir uns entschieden, da wir Algorithmen finden wollten die in einem isolierten system gute vorhersagen treffen können. Diese Algorithmen dann noch um einen Teil zu erweitern der die Nachrichten durchsucht, wäre ein weiterer Schritt um die Algorithmen noch besser zu machen, und im Nachhinein nicht sonderlich kompliziert.

All unsere Testergebnisse werden wir später in dem Paper vorstellen. Außerdem werden wir aufzeigen mit welchen Parametern wir jeweils die besten Ergebnisse erzielen konnten.

\subsection{Document overview}
Kapitel \ref{essentials} geht kurz auf die Hintergründe unseres Projekts ein und erläutert all die Dinge die ein Leser wissen muss um unsere Untersuchungen zu verstehen.

Kapitel \ref{algorithm} Beschreibt die von uns genutzten Algorithmen.

In Kapitel \ref{results} gibt einen Überblick über die Testergebnisse die wir mit den Algorithmen erzielt haben. Außerdem werden hier die verschiedenen Parameter besprochen mit denen wir die besten Ergebnisse erreicht haben.

In Sektion \ref{conclusion} werden die Resultate analysiert und deren Bedeutung für den Aktienmarkt erläutert.

In Kapitel \ref{futurework} werden wir beschreiben wie die verwendeten Algorithmen in der Zukunft noch verbessert und erweitert werden könnten.
